rgbif: R client for working with GBIF species occurrence data

### 2016-03-11 manuscript.Rmd
manuscrpt changes

---
layout: review, 11pt
linenumbers: true
title: "rgbif: R client for working with GBIF species occurrence data"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      University of California, Berkeley, CA, USA
abstract: |
      1. xxx

@@

Perhaps the most fundamental element in many fields of ecology is
the individual. The number of individuals of each species in a given
location forms the basis for many sub-fields of ecology and evolution.
Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

@@

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
http://www.gbif.org) is the largest collection of biodiversity
records globally, currently with 640 million records, 1.6 million
taxa, 15,000 datasets from 780 publishers (current as of 2015-11-25).
Many large biodiversity warehouses such as iNaturalist
(http://www.inaturalist.org), VertNet (http://vertnet.org), and USGS's
Biodiversity Information Serving Our Nation (BISON; http://bison.usgs.ornl.gov)
all feed into GBIF.

Herein, we describe the `rgbif` software library [@rgbif] for working with
GBIF data in the R programming environment [@R]. R is a widely used
language in academia, and in non-profit and private sectors.
Importantly, R makes it easy to do all of the steps of the
research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client
for getting GBIF data is a powerful tool to facilitate
reproducible research.

@@

The `rgbif` package is nearly completely written in R (a small Javascript
library is included for reading well known text), uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project
(http://ropensci.org), a developer network making R software to facilitate
reproducible research.

@@

The GBIF reqistry API services are spread across five sets of functions
via the main API:

@@

Here, we seaerch for dataset metrics for a single dataset, with uuid
`ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
Formicidae) of the World.

@@

The goal of these name functions is often to settle on a taxonomic name
known to GBIF's database. This serves two purposes: 1) when referring to
a taxonomic name, you can point to a URI on the internet, and 2) you can
search for metadata on a taxon, and occurrences of that taxon in
GBIF.

@@

One of the quickest ways to search for names is using `name_suggest()`, which
does a very quick search and returns minimal data. Here, we're searching for
the query tem _Pum_, and we get back many names:

@@

GBIF provides two ways to get occurrence data: through the
`/occurrence/search` route (see `occ_search`), or via the
`/occurrence/download` route (many functions, see below).
`occ_search()` is the main funtion for the search route,
and is more appropriate for smaller data, while
`occ_download*()` functions are more appropriate for larger
data requests.

Large is of course a subjective term. When you hit a "large dataset"
will depend primarily on the size of the your data request. GBIF imposes
for any given search a limit of 200,000 records in the search service,
after which point you can't download any more records for that search.
However, you can download more records for different searches.

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches (i.e., `latitude > 50`), whereas you can not
do that in the occurrence search API. Thus, unfortnately, we couldn't
make the query interace exactly the same for both search and
download functions.

@@

* Get a single numeric count of occurrenes - `occ_count()`
* Search for occurrences - `occ_search()`
* A simplified and optimized version of `occ_search()` - `occ_data()`
* Get occurrences by occurrence identifier - `occ_get()`
* Get occurrence metadata - `occ_metadata()`

@@

The output of `occ_search()` borrows the tidy `data.frame` idea from the `dplyr`
R package, so that no matter how large the `data.frame`, the output is easily
assessed because only a few of the records (rows) are shown, only a few columns
are shown (with others shown in name only), and metadata is shown on top of
the `data.frame` to indicate data found and returned, media
records found, unique taxonomic hierarchies returned, and the query executed.

@@

Spatial search, based on well known text format, or a bounding box set of
four coordinates

@@

Get only those occurrences with spatial issues

@@

An obvious downstream use case for species occurrence data is to map the
data. `rgbif` per se is largely not concerned with making this easier,
although we do have a simple wrappar around `ggplot2` to make it easy to
get a quick plot of occurrence data. For example, here we plot 100
occurrences for _Puma concolor_.

@@

Another package, [spoccutils][spoccutils]

@@

Some of the GBIF taxonomic services are also avialable in [taxize][taxize],
an R package that focuses on getting data from taxonomic data sources on
the web. For example, with `get_gbifid()` one can get GBIF IDs used for
a set of taxonomic names - then use those IDs in other functions in `taxize`
to get additional information, like taxonomically downstream children.

@@

GBIF occurrence data is available in the R package [spocc][spocc] via `rgbif`.
`spocc` is a unified interface for fetching species occurrence data from
many sources on the web. For example, a user can collect occurrence data
from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
pacakages to clean and vizualize the data.

@@

The following are three use cases for `rgbif`: niche modelling,
spatial change in biodiversity, and distribution mapping.

### Ecological niche modelling

```{r niche-modelling, child="use-cases/niche-modelling.Rmd"}
```

@@

# Acknowledgements

This project was supported in part by the Alfred P Sloan
Foundation (Grant 2013-6-22), and the Helmsley Foundation
(Grant XXXXXX).

@@

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
[taxize]: https://github.com/ropensci/taxize
[spocc]: https://github.com/ropensci/spocc
[spoccutils]: https://github.com/ropensci/spocc

>>>>

---
layout: review, 11pt
linenumbers: true
title: "rgbif: R client for working with GBIF species occurrence data"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
  - name: Carl Boettiger
    affiliation: boettig
    email: carl(at)ropensci.org
address:
  - code: cstar
    address: |
      University of California, Berkeley, CA, USA
  - code: boettig
    address: |
      University of California, Berkeley, CA, USA
abstract: |
      1. xxx

@@

Perhaps the most fundamental element in many fields of ecology is
the individual organism. The number of individuals of each species in
a given location forms the basis for many sub-fields of ecology and
evolution. Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

@@

# Global Biodiversity Information Facility

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
[http://www.gbif.org](http://www.gbif.org)) is the largest collection of
biodiversity records globally, currently with 643 million records, 1.6 million
taxa, 15,450 datasets from 780 publishers (as of 2016-02-09).
Many large biodiversity warehouses such as iNaturalist
([http://www.inaturalist.org](http://www.inaturalist.org)), VertNet
([http://vertnet.org](http://vertnet.org)), and USGS's Biodiversity Information
Serving Our Nation (BISON; [http://bison.usgs.ornl.gov](http://bison.usgs.ornl.gov))
all feed into GBIF.

The most important data organizational level in GBIF is the individual record. In R
using `rgbif` you'll recognize this as a row in the data.frame that is returned from
`occ_search()` or `occ_data()`. Going upstream, each record is part of a dataset,
where each dataset is submitted by an organization, organizations are organized into
nodes, datasets are published through institutions (which may be hosted at another
organization), and a network is a group of datasets (managed by GBIF).

Each record has some taxonomic name associated with it, which itself is linked to
a lot of other taxonomic data. GBIF maintains their own taxonomic data.

All this organization matters because you can navigate through GBIF data through all
these organizational levels.

Herein, we describe the `rgbif` software package [@rgbif] for working with
GBIF data in the R programming environment [@R]. R is a widely used
language in academia, as well as non-profit and private sectors.
Importantly, R makes it easy to execute all steps of the
research process, including data management, data manipulation
and cleaning, statistics, and visualization. Thus, an R client
for getting GBIF data is a powerful tool to facilitate
reproducible research.

@@

The `rgbif` package is nearly completely written in R (a small Javascript
library is included for reading well known text [@wkt]), uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project
(http://ropensci.org), a developer network making R software to facilitate
reproducible research.

@@

The GBIF registry API services are spread across five sets of functions
via the main API:

@@

Here, we search for dataset metrics for a single dataset, with uuid
`ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
Formicidae) of the World.

@@

The goal of these name functions is often to settle on a taxonomic name
known to GBIF's database. This serves two purposes: 1) when referring to
a taxonomic name, you can point to a URI on the Internet, and 2) you can
search for metadata on a taxon, and occurrences of that taxon in
GBIF.

@@

One of the quickest ways to search for names is using `name_suggest()`, which
does a very quick search and returns minimal data. Here, we're searching for
the query term _Pum_, and we get back many names:

@@

GBIF provides two ways to get occurrence data: through the
`/occurrence/search` route (see `occ_search`), or via the
`/occurrence/download` route (many functions, see below).
`occ_search()` is the main function for the search route,
and is more appropriate for smaller data, while
`occ_download*()` functions are more appropriate for larger
data requests.

GBIF imposes for any given search a limit of 200,000 records in the search
service, after which point you can't download any more records for that
search. However, you can download more records for different searches.

@@

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches (i.e., `latitude > 50`), whereas you cannot
do that in the occurrence search API. Thus, unfortunately, we couldn't
make the query interface exactly the same for both search and
download functions.

@@

* Get a single numeric count of occurrences - `occ_count()`
* Search for occurrences - `occ_search()`
* A simplified and optimized version of `occ_search()` - `occ_data()`
* Get occurrences by occurrence identifier - `occ_get()`
* Get occurrence metadata - `occ_metadata()`

@@

The output of `occ_search()` presents a compact `data.frame` so that no matter
how large the `data.frame`, the output is easily assessed because only a few of
the records (rows) are shown, only a few columns are shown (with others shown in
name only), and metadata is shown on top of the `data.frame` to indicate data
found and returned, media records found, unique taxonomic hierarchies returned,
and the query executed.

@@

Spatial search, based on well known text format [@wkt], or a bounding box set of
four coordinates

@@

Get only those occurrences with spatial issues. Spatial issues are a set of issues
that are returned in the `issues` field. They each indicate something different
about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
that the interpreted occurrence coordinates fall outside of the indicated country.
You can see how that might be useful when it comes to cleaning your data prior
to analysis/visualization.

@@

An obvious downstream use case for species occurrence data is to map the
data. `rgbif` per se is largely not concerned with making this easier,
although we do have a simple wrapper around `ggplot2` to make it easy to
get a quick plot of occurrence data. For example, here we plot 100
occurrences for _Puma concolor_.

@@

Another package, [mapr][mapr], is the perfect mapping companion to `rgbif`.
It has convenient functions for handling input data from `rgbif`, `spocc`,
or arbitrary data.frame's, and output plots for base plots, `ggplot2`,
`ggmap` (`ggplot2` with map layers underneath), and interactive maps on
GitHub gists or with Leaflet.js.

@@

Some of the GBIF taxonomic services are also available in [taxize][taxize],
an R package that focuses on getting data from taxonomic data sources on
the web. For example, with `get_gbifid()` one can get GBIF IDs used for
a set of taxonomic names - then use those IDs in other functions in `taxize`
to get additional information, like taxonomically downstream children.

@@

GBIF occurrence data is available in the R package [spocc][spocc] via `rgbif`.
`spocc` is a unified interface for fetching species occurrence data from
many sources on the web. For example, a user can collect occurrence data
from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
packages to clean and visualize the data.

@@

The following are three use cases for `rgbif`: niche modeling,
spatial change in biodiversity, and distribution mapping.

### Ecological niche modeling

```{r niche-modeling, child="use-cases/niche-modelling.Rmd"}
```

@@

# Acknowledgments

This project was supported in part by the Alfred P Sloan Foundation
(Grant No. G-2014-13485), and in part by the Helmsley Foundation
(Grant No. 2016PG-BRI004).

@@

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
[taxize]: https://github.com/ropensci/taxize
[spocc]: https://github.com/ropensci/spocc
[mapr]: https://github.com/ropensci/mapr

### 2016-03-11 manuscript.tex
manuscrpt changes

  \title{rgbif: R client for working with GBIF species occurrence data}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
      \address[cstar]{University of California, Berkeley, CA, USA}

@@

Herein, we describe the \texttt{rgbif} software library (Chamberlain et
al.) for working with GBIF data in the R programming environment (R Core
Team, 2014). R is a widely used language in academia, and in non-profit
and private sectors. Importantly, R makes it easy to do all of the steps
of the research process, including data management, data manipulation
and cleaning, statistics, and vizualization. Thus, an R client for
getting GBIF data is a powerful tool to facilitate reproducible
research.

@@

The \texttt{rgbif} package is nearly completely written in R (a small
Javascript library is included for reading well known text), uses an
\href{http://choosealicense.com/licenses/mit/}{MIT license} to maximize
use everywhere. \texttt{rgbif} is developed publicly on GitHub at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{rgbif} can be
installed from
\href{https://cran.rstudio.com/web/packages/rgbif/}{CRAN}, the
distribution network for R packages. \texttt{rgbif} is part of the
rOpenSci project (\url{http://ropensci.org}), a developer network making
R software to facilitate reproducible research.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2580).

Here, we seaerch for dataset metrics for a single dataset, with uuid
\texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
(Hymenoptera: Formicidae) of the World.

@@

One of the quickest ways to search for names is using
\texttt{name\_suggest()}, which does a very quick search and returns
minimal data. Here, we're searching for the query tem \emph{Pum}, and we
get back many names:

@@

GBIF provides two ways to get occurrence data: through the
\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
\texttt{/occurrence/download} route (many functions, see below).
\texttt{occ\_search()} is the main funtion for the search route, and is
more appropriate for smaller data, while \texttt{occ\_download*()}
functions are more appropriate for larger data requests.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
not do that in the occurrence search API. Thus, unfortnately, we
couldn't make the query interace exactly the same for both search and
download functions.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrenes - \texttt{occ\_count()}
\item
  Search for occurrences - \texttt{occ\_search()}
\item
  A simplified and optimized version of \texttt{occ\_search()} -
  \texttt{occ\_data()}
\item
  Get occurrences by occurrence identifier - \texttt{occ\_get()}
\item
  Get occurrence metadata - \texttt{occ\_metadata()}
\end{itemize}

@@

Spatial search, based on well known text format, or a bounding box set
of four coordinates

@@

An obvious downstream use case for species occurrence data is to map the
data. \texttt{rgbif} per se is largely not concerned with making this
easier, although we do have a simple wrappar around \texttt{ggplot2} to
make it easy to get a quick plot of occurrence data. For example, here
we plot 100 occurrences for \emph{Puma concolor}.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}
\caption{}
\end{figure}

Another package, \href{https://github.com/ropensci/spocc}{spoccutils}

@@

Some of the GBIF taxonomic services are also avialable in
\href{https://github.com/ropensci/taxize}{taxize}, an R package that
focuses on getting data from taxonomic data sources on the web. For
example, with \texttt{get\_gbifid()} one can get GBIF IDs used for a set
of taxonomic names - then use those IDs in other functions in
\texttt{taxize} to get additional information, like taxonomically
downstream children.

@@

GBIF occurrence data is available in the R package
\href{https://github.com/ropensci/spocc}{spocc} via \texttt{rgbif}.
\texttt{spocc} is a unified interface for fetching species occurrence
data from many sources on the web. For example, a user can collect
occurrence data from GBIF, iDigBio, and iNaturalist, and easily combine
them, then use other pacakages to clean and vizualize the data.

@@

The following are three use cases for \texttt{rgbif}: niche modelling,
spatial change in biodiversity, and distribution mapping.

\subsubsection{Ecological niche
modelling}\label{ecological-niche-modelling}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}
\caption{}
\end{figure}

@@

\section{Acknowledgements}\label{acknowledgements}

@@

>>>>

  \title{rgbif: R client for working with GBIF species occurrence data}
    \author[cstar]{Scott Chamberlain\corref{c1}}
   \ead{scott(at)ropensci.org} 
   \cortext[c1]{Corresponding author}
    \author[boettig]{Carl Boettiger}
   \ead{carl(at)ropensci.org}

      \address[cstar]{University of California, Berkeley, CA, USA}    
    \address[boettig]{University of California, Berkeley, CA, USA}

@@

Herein, we describe the \texttt{rgbif} software library (Chamberlain et
al.) for working with GBIF data in the R programming environment (R Core
Team, 2014). R is a widely used language in academia, and in non-profit
and private sectors. Importantly, R makes it easy to do all of the steps
of the research process, including data management, data manipulation
and cleaning, statistics, and visualization. Thus, an R client for
getting GBIF data is a powerful tool to facilitate reproducible
research.

@@

The \texttt{rgbif} package is nearly completely written in R (a small
Javascript library is included for reading well known text (Herring,
2011)), uses an \href{http://choosealicense.com/licenses/mit/}{MIT
license} to maximize use everywhere. \texttt{rgbif} is developed
publicly on GitHub at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
where development versions of the package can be installed, and bugs and
feature requests reported. Stable versions of \texttt{rgbif} can be
installed from
\href{https://cran.rstudio.com/web/packages/rgbif/}{CRAN}, the
distribution network for R packages. \texttt{rgbif} is part of the
rOpenSci project (\url{http://ropensci.org}), a developer network making
R software to facilitate reproducible research.

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2637).

Here, we search for dataset metrics for a single dataset, with uuid
\texttt{ec93a739-1681-4b04-b62f-3a687127a17f}, a checklist of the ants
(Hymenoptera: Formicidae) of the World.

@@

One of the quickest ways to search for names is using
\texttt{name\_suggest()}, which does a very quick search and returns
minimal data. Here, we're searching for the query term \emph{Pum}, and
we get back many names:

@@

GBIF provides two ways to get occurrence data: through the
\texttt{/occurrence/search} route (see \texttt{occ\_search}), or via the
\texttt{/occurrence/download} route (many functions, see below).
\texttt{occ\_search()} is the main function for the search route, and is
more appropriate for smaller data, while \texttt{occ\_download*()}
functions are more appropriate for larger data requests.

@@

The download API syntax is similar to the occurrence search API in that
the same parameters are used, but the way in which the query is defined
is different. For example, in the download API you can do greater than
searches (i.e., \texttt{latitude \textgreater{} 50}), whereas you can
not do that in the occurrence search API. Thus, unfortunately, we
couldn't make the query interface exactly the same for both search and
download functions.

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrences - \texttt{occ\_count()}
\item
  Search for occurrences - \texttt{occ\_search()}
\item
  A simplified and optimized version of \texttt{occ\_search()} -
  \texttt{occ\_data()}
\item
  Get occurrences by occurrence identifier - \texttt{occ\_get()}
\item
  Get occurrence metadata - \texttt{occ\_metadata()}
\end{itemize}

@@

Spatial search, based on well known text format (Herring, 2011), or a
bounding box set of four coordinates

@@

An obvious downstream use case for species occurrence data is to map the
data. \texttt{rgbif} per se is largely not concerned with making this
easier, although we do have a simple wrapper around \texttt{ggplot2} to
make it easy to get a quick plot of occurrence data. For example, here
we plot 100 occurrences for \emph{Puma concolor}.

@@

\includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}

Another package, \href{https://github.com/ropensci/mapr}{mapr}, is the
perfect mapping companion to \texttt{rgbif}. It has convenient functions
for handling input data from \texttt{rgbif}, \texttt{spocc}, or
arbitrary data.frame's, and output plots for base plots,
\texttt{ggplot2}, \texttt{ggmap} (\texttt{ggplot2} with map layers
underneath), and interactive maps on GitHub gists or with Leaflet.js.

@@

Some of the GBIF taxonomic services are also available in
\href{https://github.com/ropensci/taxize}{taxize}, an R package that
focuses on getting data from taxonomic data sources on the web. For
example, with \texttt{get\_gbifid()} one can get GBIF IDs used for a set
of taxonomic names - then use those IDs in other functions in
\texttt{taxize} to get additional information, like taxonomically
downstream children.

@@

GBIF occurrence data is available in the R package
\href{https://github.com/ropensci/spocc}{spocc} via \texttt{rgbif}.
\texttt{spocc} is a unified interface for fetching species occurrence
data from many sources on the web. For example, a user can collect
occurrence data from GBIF, iDigBio, and iNaturalist, and easily combine
them, then use other packages to clean and visualize the data.

@@

The following are three use cases for \texttt{rgbif}: niche modeling,
spatial change in biodiversity, and distribution mapping.

\subsubsection{Ecological niche
modeling}\label{ecological-niche-modeling}

@@

\includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}

@@

\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}

@@

\includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}

@@

\section{Acknowledgments}\label{acknowledgments}

@@

Herring J. 2011. OpenGIS implementation standard for geographic
information-simple feature access-part 1: Common architecture. \emph{OGC
Document} 4:122--127.
